Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-07-16 doc/paper-lcr.tex
magnitudes and radiu

Given the halo mass catalogue from the Millenium Simulation, and convergence maps produced by
ray-tracing through the MS's n-body particles, 
it is possible to test the validity of using a truncated NFW halo model to estimate 
the external convergence along any given line of sight.

Lines of sight are then chosen from random positions in the catalogue, and for each line of
sight we reconstruct the convergences from every halo within 5 arcminutes and compare the reconstructed
kappas with those from the ray tracing results. The standard deviation on  
shown in Figure \ref{fig:ScattervsTruncation}. It is clear from Figure \ref{fig:ScattervsTruncation} 
that as long as the truncation is greater than a few virial radii, the choice of truncation
radius does not broadly affect the scatter on $\kapparay$-$kapparec$; from now on we choose 
to truncate the halos at ten times their virial radius.

@@

>>>>

Given the halo mass catalogue from the Millenium Simulation, and convergence maps produced by
ray-tracing through the MS's n-body particles, 
it is possible to test the validity of the prescription outlined in Section \ref{sec:model}. We do this
using $\kapparay$ as a proxy for $\kappatrue$. The reconstruction prescription is tested given perfect
knowledge of the halos' virial masses and redshifts; this provides us with a measure of the intrinsic
errors induced by treating matter as truncated NFW halos. To quantify this 10000 lines of sight are chosen
from random positions in the catalogue, and for each line of sight we reconstruct the
convergences from every halo within 5 arcminutes and compare the reconstructed
kappas with those from the ray tracing results. The standard deviation on $\kapparay$-$\kapparec$ 
shown in Figure \ref{fig:ScattervsTruncation}. It is clear from Figure \ref{fig:ScattervsTruncation} 
that as long as the truncation is greater than a few virial radii, the choice of truncation
radius does not broadly affect the scatter on $\kapparay-\kapparec$; from now on we choose 
to truncate the halos at $10 \Rhalo$. Since $\sigma\left(\kapparay-\kapparec\right)$ does not go to
zero, it is clear that out halo based prescription does not perfectly replicate the n-body results, we refer to this
 as the intrinsic error of the prescription. The intrinsic error has a standard deviation of $\sigi \sim 0.018$, although
this figure is inflated by a few outliers in the tails of $\pr\left(\kapparay-\kapparec\right)$. Whilst we do not
focus on shear in this work, such outliers would likely be rejected by their reconstructed shear not matching the
shear of the lens model (see \citet{wong} for the practicalities of this), this makes our estimate of the intrinsic error pessimistic.

\tom{this paragraph is phrased badly}
If $\kapparec$ is to be used in a real survey it is important that it is not a biased estimator of the true $\kappax$. 
Ideally one wants $\langle \kapparec-\kappatrue \rangle$ to equal 0 and $\kapparec-\kappatrue$ to be independant of
$\kapparec$., since the void correction ensures $\langle \kapparec \rangle = 0$ any bias in $\langle \kapparec-\kapparay \rangle$ comes
from the variations in $\kapparay$. With 8.2$\times 10^8$ pixels, uniformly spaced over 196 deg$^2$ of simulated sky,
 $\langle \kapparay \rangle$ = -0.00014. However, it is improbable that such a large number of lines of sight
 could be reconstructed in reality and the mean of $\kappatrue$ for a small subset of sightlines will not necessarily
equal zero. We find this error to be
\be\label{eq:sigmav}
\sigma(\kappav) \simeq 0.036 / N^{\frac{1}{2}},
\ee
 where $N$ is the number of lines of sight, although $\pr(\kappav)$ has positive skew for small $N$.
 Care must be taken that the callibration lines of sight are not biased in $\kappax$ or the final $\pr(\kappav)$ will also be biased. For the rest of this work we use 1000  randomly chosen calibration lines of sight to calculate $\kappav$, introducing an error on $\kappax$ of $0.001$. 1000 callibration lines of sight is unfeasable currently, but should be reasonable in the LSST era. Since the true $\langle \kapparay \rangle$ = -0.00014 is two orders of magnitude smaller than intrinsic error of the halo prescription, $\kapparec$ can be seen as unbiased given a large enough sample of lines of sight. For an individual line of sight $\kapparec$ can also be shown
\begin{figure}\label{fig:magcut}
\includegraphics[width=\columnwidth]{mag_scatter.eps}
\end{figure}
\begin{figure}\label{fig:radcut}
\includegraphics[width=\columnwidth]{radius_scatter.eps}
\end{figure}
The error from the void correction, and the intrinsic error from the halo based prescription are unavoidable errors, even with perfect knowledge of halo mass and position. Before investigating the errors caused by imperfect knowledge of halo mass and redshift, we investigate the errors induced by a magnitude limited reconstruction and a field of view limited reconstruction. The halos in our catalogue are given magnitudes by the semi-analytic model of \citet{SAM} and by applying magnitude cuts to our catalogue, we can investigate the scatter caused by unobserved halos. The majority of the convergence comes from halos with an $i$ magnitude between 18 and 24. Figure \ref{fig:magcut} shows that the width of $\kapparec-\kapparay$ decreases quickly between $i=18$ and $i=24$. Objects brighter than $i=18$ are either too rare or too close to the observer to make a significant contribution to the convergence. Objects fainter than $i=24$ are too small to be important unless they are extremely close to the line of sight where it is likely that neglecting stellar mass and using a spherical NFW prescription are too naive to adequately reconstruct $\kappax$ - it is likely that these ultra-faint halos (or substructures) are the major source of the scatter in $\kapparec-\kapparay$, but a deeper survey will not be sufficient to decrease the size of this scatter. Since a large number of callibration lines of sight are needed to shrink the width of $\pr(\kappav)$, it may prove unrealistic to reconstruct the calibration lines of sight out to 5 arcminutes. \citet{Auger2007} showed that the majority significant perturbers are within 15 arcseconds of the lens. Figure \ref{fig:radcut} shows the width of $\kapparec-\kapparay$ as a function of reconstruction radius; in almost all cases there is no significant contribution to $\kappa$ from halos that are centred more than 1 arcminute away, although large groups or clusters can sometimes still make a contribution as far out as four arcminutes. For the rest of this work we will continue to model all of the halos out to 5 arminutes with an $i$ magnitude greater than 26, although a reconstruction with fields going out to two arcminutes and down to $i<24$ would do almost as well.

@@

Emphasize that this can be done before monitoring time is invested.
